---
title: "FInancial time series model: Returns from Beef"
author: "hillary"
date: "29 January 2019"
output: html_document
---
Load data
```{r}
beef_price=read.csv("beef_price.csv",header=T)
head(beef_price)
```
convert to time series

```{r}
Beefprice=ts(beef_price,frequency = 12,start = 2010,end=2018)

```
just check to confirm its a time series

```{r}
str(Beefprice)
```

Package installation
install the tseries package to perform stationarity test
```{r}
library(tseries)
```
FINANCIAL TIME SERIES ANALYSIS TO ANALYSE THE TREND OF RETURNS 
RETURNS(Rt)=P(t)/P(t-1)
log(Rt)=log(Pt)-log(Pt-1)

DECOMPOSING THE SERIES TO GET RID OF SEASONALITY
Get the log returns with lag=16 years back to remove the multiplicative seasonality
```{r}
log_returns <- diff(log(Beefprice), lag=10)
head(log_returns)
```


Test for stationarity  if alpha<0.05 reject null hypothesis
```{r}
adf.test(log_returns)
```

```{r}
plot(log_returns)
```

Plot the ACF and PACF to determine the order of the ARIMA model
```{r}
acf(log_returns)
```
```{r}
pacf(log_returns)
```
From the ACF 2 lags cut off the dotted line hence q=2 as for p=3 coz 3 lags cut off the dotted line in the PACF order=(3,1,3)=(p,d,q), d=2 coz difference was done twice.

ARIMA ON THE LOG RETURNS

```{r}
mymodel=arima(log_returns,order = c(2,2,5))
mymodel

```

FORECASTING FOR THE YEAR 2020
```{r}
log_returns_2019<-predict(mymodel,n.ahead=12)
log_returns_2019




