---
title: "Naive Bayes markdown"
author: "hillary"
date: "28 January 2019"
output: html_document
---
---
Classification method that uses probability to estimate the likelihood that an observation falls into certain categories
Classifiers based on Bayesian methods utilize training data to calculate an observed probability of each outcome based on the evidence provided by feature values. When
the classifier is later applied to unlabeled data, it uses the observed probabilities to predict the most likely class for the new features.
It's a simple idea, but it results in a method that often has results on par with more sophisticated algorithms

DATA LOADING, INSTALL PACKAGES & DESCREPTIVE STATS
---

```{r}
sms_raw=read.csv('sms_raw.csv',stringsAsFactors = FALSE)
head(sms_raw)
#install.packages("tm")#text mining packages 
#install.packages("SnowballC")
#install.packages("gmodels")
```

```{r}
as.character(sms_raw$text[1])
str(sms_raw$text)
table(sms_raw$type)
```


CLEANING AND STANDARDIZING THE DATA
```{r}
library(tm)
sms_corpus <- VCorpus(VectorSource(sms_raw$text))
sms_corpus
print(sms_corpus)
inspect(sms_corpus[1:2])
as.character(sms_corpus[[1]])
lapply(sms_corpus[1:2], as.character)

```

Convert to lower case
```{r}
sms_corpus_clean <- tm_map(sms_corpus,content_transformer(tolower))
as.character(sms_corpus_clean[[1]])
```
Remove numbers
```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removeNumbers)
as.character(sms_corpus_clean[[1]])
```
Remove stopwords
```{r}
sms_corpus_clean<-tm_map(sms_corpus_clean,removeWords,stopwords())
as.character(sms_corpus_clean[[1]])
```
Remove punctuation
```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removePunctuation)
as.character(sms_corpus_clean[[1]])
```
Stem the document; returns the same vector of terms in its root form
```{r}
library(SnowballC)
sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
as.character(sms_corpus_clean[[1]])
```
Strip whitespace
```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, stripWhitespace)
as.character(sms_corpus_clean[[1]])
```
DATA PREPARATION  splitting text documents into words
1. Creating a DTM sparse matrix, given a tm corpus, involves a single command:
```{r}
sms_dtm <- DocumentTermMatrix(sms_corpus_clean)
str(sms_dtm)
rowTotals<-apply(sms_dtm,1,sum)#Find the sum of words in each Document
sms_dtm.new   <- sms_dtm[rowTotals> 0, ] #remove all docs without words
```
if the two sms_dtm are not similar then stopwords must be the problem thus replace it with the following function to correct it.
```{r}
stopwords = function(x) { removeWords(x, stopwords()) }
```
2. Creating training and test datasets
```{r}
sms_dtm_train <- sms_dtm.new[1:4169, ]
sms_dtm_test <- sms_dtm.new[4170:7028, ]
sms_train_labels <- sms_raw[1:4169, ]$type
sms_test_labels <- sms_raw[4170:7028, ]$type
#To confirm that the subsets are representative of the complete set of SMS data, let's compare the proportion of spam in the training and test data frames:
prop.table(table(sms_train_labels))
prop.table(table(sms_test_labels))
```
3. Visualizing text data  word clouds
```{r}
library("RColorBrewer")
library("wordcloud")
wordcloud(sms_corpus_clean, min.freq = 50, random.order = FALSE)
```
split for separate visualisations
```{r}
spam <- subset(sms_raw, type == "spam")
ham <- subset(sms_raw, type == "ham")
wordcloud(spam$text, max.words = 40, scale = c(3, 0.5))
wordcloud(ham$text, max.words = 40, scale = c(3, 0.5))
```
4. Creating indicator features for frequent words
```{r}
findFreqTerms(sms_dtm_train, 5)
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
str(sms_freq_words)
```
4. Creating indicator features for frequent words
```{r}
findFreqTerms(sms_dtm_train, 5)
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
str(sms_freq_words)
```
Create train and test for freq words
```{r}
sms_dtm_freq_train<- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]
convert_counts <- function(x) {
x <- ifelse(x > 0, "Yes", "No")
}
```
Apply the convert function to create the train and test 
```{r}
sms_train <- apply(sms_dtm_freq_train, MARGIN = 2,convert_counts)
sms_test <- apply(sms_dtm_freq_test, MARGIN = 2,convert_counts)
```
TRAINING A MODEL ON THE DATA
```{r}
str(sms_train_labels)
head(sms_train_labels)
library("naivebayes")
sms_classifier <- naive_bayes(sms_train, sms_train_labels)
```
EVALUATING MODEL PERFORMANCE
Prediction
```{r}
sms_test_pred <- predict(sms_classifier,sms_test)
#Check the false posiutives and negatives
library(gmodels)
CrossTable(sms_test_pred, sms_test_labels,prop.chisq = FALSE, prop.t = FALSE,dnn = c('predicted', 'actual'))
```








